### 2015-11-12 thesis-proposal/problem.tex
Updated timeline for proposal

The second form in the syntax of $\hat{e}$ that has no corresponding form in the syntax of $e$ is the form for TSM application to a delimited body, $\eta\,\texttt{/}b\texttt{/}$. Again because substitution for TSM variables is performed statically, we only need a rule for the case where $\eta$ is a TSM definition:
\begin{mathpar}
\inferrule[T-U-TSM-ap]{
	\Delta \vdash \tau~\mathtt{type}\\
	\emptyset~\emptyset \vdash \hat{e}_\text{parse} \leadsto e_\text{parse} : \mathsf{Body} \rightharpoonup \mathsf{ExpOption}\\\\
	b \downarrow e_\text{body} : \mathsf{Body}\\
	e_\text{parse}(e_\text{body}) \Downarrow \mathsf{Some}(e_\text{exp})\\
	e_\text{exp} : \mathsf{Exp} \uparrow \dot{e}_\text{exp}\\\\
	\Delta~\Gamma; \emptyset~\emptyset \vdash \dot{e}_\text{exp} \leadsto e : \tau
}{
	\Delta~\Gamma \vdash \mathsf{syntax}~@~\tau~\{\hat{e}_\text{parse}\}~\texttt{/}b\texttt{/} \leadsto e : \tau
}
\end{mathpar}
The premises can be understood as follows, in order:
\begin{enumerate}
\item The first premise checks that the type specified by the TSM is valid.
\item The second premise typechecks and expands the parse function.
\item The third premise encodes the body, $b$, as a term, $e_\text{body}$ of type $\mathsf{Body}$ (we will give more details on how bodies are encoded in the dissertation, but omit the definition of $\mathsf{Body}$ here for concision).
\item The fourth premise applies the expanded parse function to the encoding of the body to produce an encoding of the expansion, $e_\text{exp}$, of  type $\mathsf{Exp}$. 
\item Values of type $\mathsf{Exp}$ map onto \emph{marked expressions}, $\dot{e}$. Per Figure \ref{fig:lambda-tsm-syntax}, marked expressions can contain variables, type variables and marked types, $\dot{\tau}$, so there is also a mapping from values of types abbreviated $\mathsf{Var}$, $\mathsf{TVar}$ and $\mathsf{Type}$ onto variables, type variables, and {marked types}, respectively. We also omit the full definitions of these types for concision (cf. the SML/NJ Visible Compiler library \cite{SML/VisibleCompiler} for an example of such an encoding of the abstract syntax of a language). These mappings are specified by the \emph{expansion decoding judgements}:

@@

The fifth premise decodes $e_\text{exp}$ to produce the \emph{marked expansion}, $\dot{e}_\text{exp}$. 
\item The final premise valides the expansion by checking the marked expansion against the type specified by the TSM, and generates the final expansion, $e$, according to the rules defining the \emph{expansion validation judgements}:

@@

ML does not presently support pattern matching over values of an abstract data type. However, there have been proposals for adding support for pattern matching over abstract data types defined by modules having a ``datatype-like'' shape, e.g. those that define a case analysis function like the one specified by \lstinline{RX}, shown in Sec. \ref{sec:examples}. We leave further discussion of such a facility and of parameterized TPSMs also as remaining work (see Sec. \ref{sec:syntax-timeline}).

@@

In the dissertation, I plan to first present the formal system $\mathcal{L}^\text{TSM}$. The account in Sec. \ref{sec:tsm-formalism} represents the first and most substantial step of this plan. This is complete on paper, so I only need to typeset the omitted details, including the metatheory. I plan to finish this by the \textbf{end of October}. This system will serve as the basis for several extensions that capture other constructs described above, so I will ask the committee to read this section in detail once it is complete.

The first extension will add support for simple TSLs over unparameterized datatypes. This has already been written up in \cite{TSLs}, so I plan to have this in final form in the \textbf{first week of November}.

The next extension will add support for pattern matching over these datatypes, and TPSMs. I will base this extension on the account of pattern matching given in \emph{PFPL} \cite{pfpl}. I plan to have a rough sketch of how this should work by \textbf{late-November}, in consultation with the committee, and a final write-up of it by \textbf{mid-December}.

The final extension will add support for type and module parameters. To do so, I will assume that the available module paths and their signatures are being tracked by a module context in the static semantics, without detailing how the module language populates this context (i.e. I will only specify the ``core language''). I plan to sketch this out and discuss it with my committee also \textbf{during November} and finish writing the details up by the \textbf{last week of December}.

Over the course of the \textbf{Spring semester}, I plan to write the remaining sections (e.g. additional background, related and future work sections, and additional examples), flesh out the proofs of minor lemmas and respond to the committee's requests for edits. Based on this timeline, I plan to defend before the \textbf{end of the Spring semester}.

My current draft of the dissertation document will be available on GitHub:

>>>>

The second form in the syntax of $\hat{e}$ that has no corresponding form in the syntax of $e$ is the form for TSM application to a delimited body, $\eta\,\texttt{/}b\texttt{/}$. Again because substitution for TSM variables is performed statically, we only need a rule for the case where $\eta$ is a TSM definition:
\begin{mathpar}
\inferrule[T-U-TSM-ap]{
	\Delta \vdash \tau~\mathtt{type}\\
	\emptyset~\emptyset \vdash \hat{e}_\text{parse} \leadsto e_\text{parse} : \mathsf{Body} \rightharpoonup \mathsf{ExpOption}\\\\
	b \downarrow e_\text{body} : \mathsf{Body}\\
	e_\text{parse}(e_\text{body}) \Downarrow \mathsf{Some}(e_\text{exp})\\
	e_\text{exp} : \mathsf{Exp} \uparrow \dot{e}_\text{exp}\\\\
	\Delta~\Gamma; \emptyset~\emptyset \vdash \dot{e}_\text{exp} \leadsto e : \tau
}{
	\Delta~\Gamma \vdash \mathsf{syntax}~@~\tau~\{\hat{e}_\text{parse}\}~\texttt{/}b\texttt{/} \leadsto e : \tau
}
\end{mathpar}
The premises can be understood as follows, in order:
\begin{enumerate}
\item The first premise checks that the type specified by the TSM is valid.
\item The second premise typechecks and expands the parse function.
\item The third premise encodes the body, $b$, as a term, $e_\text{body}$ of type $\mathsf{Body}$ (we will give more details on how bodies are encoded in the dissertation, but omit the definition of $\mathsf{Body}$ here for concision).
\item The fourth premise applies the expanded parse function to the encoding of the body to produce an encoding of the expansion, $e_\text{exp}$, of  type $\mathsf{Exp}$. 
\item The fifth premise decodes $e_\text{exp}$ to produce the \emph{marked expansion}, $\dot{e}_\text{exp}$. Every value of type $\mathsf{Exp}$ corresponds to a \emph{marked expressions}, $\dot{e}$. Per Figure \ref{fig:lambda-tsm-syntax}, marked expressions can contain variables, type variables and marked types, $\dot{\tau}$, so there is also a mapping from values of types abbreviated $\mathsf{Var}$, $\mathsf{TVar}$ and $\mathsf{Type}$ onto variables, type variables, and {marked types}, respectively. We also omit the full definitions of these types for concision (cf. the SML/NJ Visible Compiler library \cite{SML/VisibleCompiler} for an example of such an encoding of the abstract syntax of a language). These mappings are specified by the \emph{expansion decoding judgements}:

@@

\item The final premise valides the expansion by checking the marked expansion against the type specified by the TSM, and generates the final expansion, $e$, according to the rules defining the \emph{expansion validation judgements}:

@@

ML does not presently support pattern matching over values of an abstract data type. However, there have been proposals for adding support for pattern matching over abstract data types defined by modules having a ``datatype-like'' shape, e.g. those that define a case analysis function like the one specified by \lstinline{RX}, shown in Sec. \ref{sec:examples}. We leave further discussion of such a facility and of parameterized TPSMs as remaining work (see Sec. \ref{sec:syntax-timeline}).

@@

In the dissertation, I plan to first describe the simplest style of TSMs (i.e. unparameterized expression TSMs) by example. This is \textbf{complete}. I will next present a formal system that formalizes such TSMs. The account in Sec. \ref{sec:tsm-formalism} represents the first and most substantial step of this plan. This is complete on paper, so I mainly need to typeset the omitted details, including the metatheory. I have made substantial progress as of this writing (see link below) and I plan to finish this by the \textbf{middle of November}. This system will serve as the basis for several extensions that capture other constructs described above, so I will ask the committee to read this section in detail once it is complete.

The first extension will add support for pattern matching. I will base this extension on the account of pattern matching given in \emph{PFPL} \cite{pfpl}. I plan to have a rough sketch of how this should work by \textbf{late-November}, in consultation with the committee, and a final write-up by Black Friday (i.e. the \textbf{third week of December}).

The next extension will add support for value, type and module parameters. To do so, I will assume that the available module paths and their signatures are being tracked by a module context in the static semantics, without detailing how the module language populates this context (i.e. I will only specify the ``core language''). I plan to sketch this out and discuss it with my committee \textbf{during November and December} and finish writing the details up by the \textbf{beginning of Spring semester}.

Beginning in the \textbf{Spring semester}, I will write up the work on type-specific languages. A variant of this work has been published already, and the groundwork for it will have already been laid down in the earlier chapters, so I anticipate being done with this by \textbf{the second week of February}. I will finish the remaining sections (e.g. additional background, related and future work sections, and additional examples), flesh out the proofs of minor lemmas and respond to the committee's requests for edits for the remainder of the semester. Based on this timeline, I plan to defend before the \textbf{end of the Spring 2016 semester}.

My current draft of the dissertation document will be available and updated on an ongoing basis on GitHub:
